Mental Health of U.S. Citizens During the COVID-19 Pandemic

### 2022-03-21 outputs/paper/paper.Rmd
final changes, spelling check and refinements

The GSS has further made changes to the survey response options as a result of the shift from in-person interviews to web and phone-based interviews. Originally the survey contains “volunteered responses” that are not displayed to the responders and is used by the interviewer. An example would be the option “about right” in response to the question “In general, do you think the courts in this area deal too harshly or not harshly enough with criminals?” ([@norc]). However, in this year's survey, the interviewers are unable to add these volunteered responses due to the lack of interaction on the web and phone-based interviews. Therefore, the GSS adopted a new structure for these volunteered responses on the web mode by creating two different versions of forms with one containing the volunteered responses as an option and the other without. Thus, any changes in the public opinion seen in the 2021 GSS data could be due to either actual changes in public opinion or the result of this newly adopted methodology and should be closely monitored and considered in analysis on the survey.

@@

```{r tab1, warning=FALSE, fig.cap="Summary of US GSS Variables", message = FALSE, echo=FALSE, fig.width=5, fig.height=5}
gss2021 <- gss2021 %>% 
  mutate(`Mental Health Score` = hlthmntl) %>%
  mutate(`Physical Health Score` = hlthphys) %>%
  mutate(`Age when first child born` = agekdbrn)
USData <- gss2021 %>% select(`Mental Health Score`, `Physical Health Score`, age, `Age when first child born` )
sumtable(USData,  title = 'US GSS Summary Statistics', out = 'kable')
```

@@

From this table, we observe the average mental health of respondents as well as their physical health as below the average of the possible responses (3 being the statistically average response), with a slightly higher average physical health. Their standard deviation, or measure of variation; is nearly identical. We also observe the summary statistics of the age the respondents were when they had their first child.

>>>>

The GSS has further made changes to the survey response options as a result of the shift from in-person interviews to web and phone-based interviews. Originally the survey contains “volunteered responses” that are not displayed to the responders and is used by the interviewer. An example would be the option “about right” in response to the question “In general, do you think the courts in this area deal too harshly or not harshly enough with criminals?” [@norc]. However, in this year's survey, the interviewers are unable to add these volunteered responses due to the lack of interaction on the web and phone-based interviews. Therefore, the GSS adopted a new structure for these volunteered responses on the web mode by creating two different versions of forms with one containing the volunteered responses as an option and the other without. Thus, any changes in the public opinion seen in the 2021 GSS data could be due to either actual changes in public opinion or the result of this newly adopted methodology and should be closely monitored and considered in analysis on the survey.

@@

```{r tab1, warning=FALSE, fig.cap="Summary of US GSS Variables", message = FALSE, echo=FALSE, fig.width=5, fig.height=5}
gss2021 <- gss2021 %>% 
  mutate(`Mental Health Score` = hlthmntl) %>%
  mutate(`Physical Health Score` = hlthphys) %>%
  mutate(`Age when first child born` = agekdbrn) %>%
  mutate(`Age` = age)
USData <- gss2021 %>% select(`Mental Health Score`, `Physical Health Score`, `Age`, `Age when first child born` )
sumtable(USData,  title = 'US GSS Summary Statistics', out = 'kable')
```

@@

From this table, we observe the average mental health of respondents as well as their physical health as below the average of the possible responses (3 being the statistically average response), with a slightly higher average physical health. Their standard deviation, or measure of variation; is nearly identical. We also observe the summary statistics of the age of the respondents, and when they had their first child.
